Covariate analysis of CMC RNASeq data reprocessed with Star-Feature counts pipeline (ACC with TMM normlaisation and Custom Design 1)

### 2016-11-07 code/Rmd/CMC_ACC_StarAligned_TMM_CD1.Rmd
Added markdown to process ACC data with custom design

---
title: "Covariate analysis of CMC RNASeq data reprocessed with Star-Feature counts pipeline (ACC with TMM normlaisation)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knitr)
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./CMC_ACC_StarAligned_TMM.Rmd",
                                 parentId ="syn7501880",
                                 entityName = 'ACC - STAR Reprocessed - with TMM')
```

@@

thisFileName <- 'CMC_ACC_StarAligned_TMM.Rmd'

@@

### Normalisation (iterative design)
Since many covariates are correlated, re-normalising COUNTS with an iterative design matrix. Here "Dx" is chosen as the primary variable of interest (i.e., covariate selection is conditioned on diagnosis)

Adding Instituion and Gender a priori to variable selection as they are correlated with Dx
```{r iterativeNorm, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Dx', 'Institution', 'Gender');

# Assign residual covariates
residualSigCovars = preAdjustedSigCovars
covariatesEffects = residualSigCovars$Effects.significantCovars[setdiff(residualSigCovars$significantCovars, postAdjustCovars)]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()

loopCount = 0 
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 100){
  writeLines(paste('Using following covariates in the model:',
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects'))

@@

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates), c(postAdjustCovars))

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                                COVARIATES[, residCovars, drop=F], 
                                                'adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)

  # Add postadjusted covariates (if any)
  covariatesEffects = residualSigCovars$Effects.significantCovars[setdiff(residualSigCovars$significantCovars, 
                                                                          c(postAdjustCovars))]

  postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects)))
  loopCount = loopCount + 1
}

modelStr <- paste(paste(gsub('_','\\\\_',postAdjustCovars), collapse=', '),
                  'as fixed effects')

tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`

### Sanity check
```{r residualSigCovars.manual, fig.width=10, fig.height=6}
residualSigCovars[["PC_res"]][[2]]$plotData
```
### Residual calculation
Calculate weighted residuals and add back "Dx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = grep("Dx", colnames(DM1$design), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "ACC - STAR Reprocessed - with TMM", parentId = parentId)
CODE <- synStore(CODE)

>>>>

---
title: "Covariate analysis of CMC RNASeq data reprocessed with Star-Feature counts pipeline (ACC with TMM normlaisation and Custom Design 1)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knitr)
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./CMC_ACC_StarAligned_TMM_CD1.Rmd",
                                 parentId ="syn7501880",
                                 entityName = 'ACC - STAR Reprocessed - with TMM - Custom Design 1')
```

@@

thisFileName <- 'CMC_ACC_StarAligned_TMM_CD1.Rmd'

@@

### Normalisation (custom design)
Normalising COUNTS with a custom design matrix. Here Instituiton, Gender, LibraryBatch, Age\\_of\\_Death, PMI, RIN, RIN2, EV.1, EV.2, EV.3, EV.4, EV.5 are adjusted conditioned on Dx
```{r customNorm, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Dx', 'Institution', 'Gender', 'LibraryBatch', 'Age_of_Death',
                     'PMI', 'RIN', 'RIN2', 'EV.1', 'EV.2', 'EV.3', 'EV.4', 'EV.5');

writeLines(paste('Using following covariates in the model:',
                 paste(postAdjustCovars, collapse=', '),
                 'as fixed effects'))

@@

# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                              COVARIATES, 
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
```

### Sanity check
```{r residualSigCovars.manual, fig.width=10, fig.height=6}
residualSigCovars[["PC_res"]][[2]]$plotData
```

### Residual calculation
Calculate weighted residuals and add back "Dx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = grep("Dx", colnames(DM1$design), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```

Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "ACC - STAR Reprocessed - with TMM - Custom Design 1", parentId = parentId)
CODE <- synStore(CODE)
